### 2020-07-14 paper.tex
added some content to involutive mcmc on dictionaries

\section{INVOLUTIVE MCMC ON GENERAL STATE SPACES}

@@

\subsection{Distributions on dictionaries}

@@

\textbf{Notation for dictionaries.} Given a dictionary $m = (A, \mathbf{x})$, we write $A_m$ for $A$ and $m[a]$ for the value $x_a$ associated with a key $a \in A$.

\textbf{Densities on finite dictionaries.} Probability distributions on dictionaries can be defined by (i) choosing 
domains $V_a$ for each address $a$ of interest, and (ii) choosing a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. We now give several examples
to build intuition.

@@

@@

\section{INVOLUTIVE MCMC} \label{sec:involutive-mcmc}
Involutive MCMC (Algorithm~\ref{alg:involutive-mcmc}) defines a transition kernel that is stationary for a probabilistic model $p$, based on an auxiliary probability kernel $q$ and a deterministic involution $f$. Different choices of $q$ and $f$ recover instances of classic algorithms from the literature (e.g., Reversible-Jump MCMC~\citep{green}, TODO: others). This section briefly reviews involutive MCMC and gives a general measure-theoretic formulation of the algorithm on arbitrary state-spaces. In the following sections, we will describe how the implementations of involutive MCMC algorithms can be automated using probabilistic and differentiable programming.

\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc-move}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary variables}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \left( \frac{d (\mu \circ f^{-1})}{d \mu} (x, y)\right)$
    % TODO check it
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC}
\label{alg:involutive-mcmc}
\end{algorithm}

@@

\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc-move}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary variables}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \left( \frac{d (\mu \circ f^{-1})}{d \mu} (x, y)\right)$
    % TODO check it
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC}
\label{alg:involutive-mcmc}
\end{algorithm}

>>>>

\section{INVOLUTIVE MCMC ON GENERAL STATE SPACES} % on traces ??

@@

\subsection{Distributions on Dictionaries}

@@

\textbf{Notation for dictionaries.} Given a dictionary $x = (A, \mathbf{x})$, we write $A_m$ for $A$ and $m[a]$ for the value $x_a$ associated with a key $a \in A$.

\textbf{Probability distributions on finite dictionaries.} 
When all keys are discrete (have countable $V_a$ and counting measure for $\mu_{a}$), a probability distribution on dictionaries is defined by a $p$ is a probability mass function $p : \mathcal{D} \to [0, 1]$ that assigns a probability $p(m)$ to each dictionary $m \in \mathcal{P}$ such that $\sum_{x \in \mathcal{D}} p(m) = 1$.
More generally a probability distribution in dictionaries is defined by a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. We now give several examples
to build intuition.

% TODO keys vs addresses

% TODO how to notate dictionaries -- 'm' or 'd' or 'x' or 't'?

@@

TODO

@@

@@

% \left( \frac{d (\mu \circ f^{-1})}{d \mu} (x, y)\right)

Suppose the set of possible keys $\mathcal{A}$ can be partitioned $\mathcal{A} = \mathcal{I} \cup \mathcal{K}$ where is a set of \emph{discrete keys} $\mathcal{I}$ that use countable $V_a$ and counting measure $\mu_a$ for $a \in \mathcal{I}$ and $\mathcal{K}$ is a set of \emph{continuous keys} that use $V_a = \mathbb{R}^{d_a}$ for some $d_a$ and Lebesgue-measure $\mu_a$ for $a \in \mathcal{K}$.
%And suppose that there exists an involution $f : Z \to Z$ such that (i) 
Suppose there is a countable partition of $Z$ into $\{Z_e : e \in E\}$ such that across each $Z_e$ the set of keys $A$ and the values $\mathbf{x}_{\mathrm{d}}$ of discrete keys are both constant (so that each set $Z_e$ is isomorphic to a Euclidean space of assignments to the continuous keys).
Suppose there is an involution $g : E \to E$ between elements of the partition (a function with $g = g^{-1}$), and a family of continuously differentiable bijections $h_{e} : Z_e \to Z_{g(e)}$ for each $e$ with $h_{e} = h_{g(e)}^{-1}$.
Let $e(z) \in E$ denote the element of the partition for an alement $z \in Z$.
Then, $f : Z \to Z$ given by $f(z) := h_{e(z)}(z)$ is an involution:
\begin{equation}
f(f(z)) = h_{e(f(z))}(h_{e(z)}(z)) = h_{g(e(z))}(h_{e(z)}(z)) = z
\end{equation}
Let $|J h_e|(z)$ denote the absolute value of the Jacobian (determinant) of $h_e$, evaluated at $z$.
Then, involutive MCMC is:
\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc-move}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary variables}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot |J h_e|(z)$
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC}
\label{alg:involutive-mcmc}
\end{algorithm}

\section{AUTOMATING INVOLUTIVE MCMC} \label{sec:involutive-mcmc}

Involutive MCMC (Algorithm~\ref{alg:involutive-mcmc}) defines a transition kernel that is stationary for a probabilistic model $p$, based on an auxiliary probability kernel $q$ and a deterministic involution $f$. Different choices of $q$ and $f$ recover instances of classic algorithms from the literature (e.g., Reversible-Jump MCMC~\citep{green}, TODO: others). This section briefly reviews involutive MCMC and gives a general measure-theoretic formulation of the algorithm on arbitrary state-spaces. In the following sections, we will describe how the implementations of involutive MCMC algorithms can be automated using probabilistic and differentiable programming.
